%! Author = jasmincapka
%! Date = 14.07.22

Imagine the following situation~\parencite[][1-2]{patreon}:
You are trying to gather public data to conduct an analysis on, but have to rely on web-scraping because the data is not directly available.
Because of this process, you cannot be sure about the completeness of the data and the validity of the resulting conclusions.
Then, because of a malicious hack, data including the needed public data for the analysis but also private data suddenly becomes directly available.

The question now arises, whether it is appropriate to use data from illicit sources for the planned research.
Therefore, this discussion focuses on the ethics of using hacked data in data science.\\
Hacked data in this context means ``data obtained in an unauthorized manner through illicit access''~\parencite[][744]{nature}.
Additionally, this paper assumes that the person potentially using the hacked data is in no way responsible for the extraction of the hacked data.

\section*{Reconstruction of problem}

This section focuses on the background, why using hacked data in data science needs ethical considerations according to~\textcite[][745]{nature}:

The ``legal and ethical boundaries of using hacked data for research are blurred''~\parencite[][745]{nature}.
As hacked data is available online, it is in the public domain which makes it per se permissible to use as the usage is not against the law.\\
Still, only because it is legally permissible, it does not account for the social responsibility of scientists and their moral obligations.
It has to be determined, whether the use of hacked data is compatible with research ethics.\\
Currently, there is few ``ethical guidance on the use of hacked data for research''~\parencite[][745]{nature}.
There are Institutional Review Boards or similar bodies such as Research Ethics Committees to ensure ethical research regarding personal data and prevent scientific misconduct.
Still, their oversight function can be bypassed.\\
There has been some effort to define best practices among researches, but the usage of hacked data has not yet been addressed directly.
This means, no general normative guidance is provided.

Therefore, the question needs to be addressed, whether, and under which conditions, research using hacked data might be morally justifiable~\parencite[][745]{nature}.

\section*{Reconstruction of arguments}

The three most relevant papers that bring arguments in favor of and against the usage of hacked data in data science are:
\textcite{patreon} examine the topic from their own perspective regarding the previously mentioned case.
\textcite{nature} cover the general debate and develop a set of guidelines for the potential usage of hacked data.
\textcite{acm} weights arguments, taking the ACM Code of Ethics into account.

The first argument in favor of the usage is that the data is already public~\parencite[][5]{patreon}.
Therefore, it can be used analogously to any other publicly accessible data~\parencite[][23]{acm}.
The illegality of its release would then be only relevant for evaluating the data quality.\\
Another argument is ``cross-domain consistency''~\parencite[][746]{nature}.
As hacked data is used in other professions like journalism, it should be usable in data science as well~\parencite[][745]{nature}.\\
Additionally, the data is of public value, meaning, an analysis can serve the public good and compensate for harms of a data breach~\parencites[][5]{patreon}[][23]{acm}.
In this case, the potential public benefits from using hacked data outweigh the harms caused by retrieving it~\parencite[][23]{acm}.\\
Furthermore, the data has a unique character as it may not be retrievable via any other conventional method in terms of accuracy and amount of data~\parencites[][745]{nature}[][5]{patreon}.\\
Lastly, even if the data could be attained otherwise, using hacked data is an ``optimization of resources''~\parencite[][746]{nature}.
Otherwise, doing the same research would cost a large amount of resources for conventionally extracting data that is already publicly available~\parencite[][745]{nature}.

In contrast to those arguments, using hacked data in data science might cause secondary harm to people, either individually or on a group level.~\parencites[][746]{nature}[23]{acm}.
The condition of hacked data being publicly available does not eliminate the potential of additional risks through further analysis.\\
Furthermore, using hacked data is a violation of the expectation of privacy of the people, the data is about~\parencite[][5]{patreon}.
There is only limited ability to differentiate between public and private, personally identifiable information within hacked data.
This makes it impossible to respect data privacy even if the data is de-identified before publication~\parencite[][746]{nature}.\\
In addition, ``scientists [are] not allowed to publish information without explicit and informed consent''~\parencite[][745]{nature}.
As consent is missing in case of hacked data, using hacked data would be using people's data without consent~\parencite[][5]{patreon}.\\
Another argument is that hacked data might be biased or of lower quality than legitimately obtained data~\parencite[][746]{nature}.
This might lead to a lowering of scientific standards regarding integrity and research ethics and can have a negative effect on the perception of science by society.
Unethical methods used to attain hacked data might mark any research using it as unethical~\parencite[][24]{acm}.\\
Additionally, the usage of hacked data is not mandatory~\parencite[][5]{patreon}.
Other data can be ethically extracted and used, even if that means not analyzing a certain topic at all or researching from a different perspective.\\
Lastly, the usage of hacked data might legitimize criminal activity as ``research''~\parencites[][5]{patreon}[][24]{acm}.
Therefore, refusing to use hacked data is an important statement about conducting research ethically and sets an example for other data scientists in the future.

\section*{General guidelines}

Considering those arguments,~\textcite{nature} concludes that ``research using hacked data should be deemed unethical, hence impermissible, by default''~\parencite[][747]{nature}, because the potential negative consequences of using hacked data in data science outweigh the benefits.
Nevertheless, they argue, that in some cases the ``potential scientific benefits of using illicitly obtained data may outweigh [\ldots] the foreseeable harms''~\parencite[][747]{nature}.
~\textcite[747-748]{nature} propose a formal research ethics assessment and formulate the following six ethical requirements as necessary for legitimately using hacked data in data science.

The first requirement is uniqueness.
This means proof needs to be provided, that the hacked data could not be retrieved via conventional methods and, thus, that they are of unique value.\\
Additionally, a risk-benefit assessment needs to be conducted.
This assessment must show, that an analysis of hacked data would be of high public value, clearly outweighing the anticipated harm, not being more than foreseeable in daily life.
The analysis must mean ``no more-than-minimal risk to the data subjects''~\parencite[][747]{nature}.\\
Furthermore, in case of personally identifiable data the scientists must obtain ``explicit and informed consent'' of the people, the data is about.\\
In addition, for traceability, a record of how and where all data have been obtained must be provided.\\
The requirement of privacy includes several aspects.
Statements regarding the assessment of personally identifiable data without consent, safeguards to prevent re-identification, information in case of insecure data transmission, and legal responsibility need to be provided.
A repository regarding the analysis of hacked data must not be made public.\\
Lastly, Institutional Review Board approval is needed.
This means that the meeting of the prior conditions should not be done via self-assessment.
Instead, an exception to the initial unethical-by-default rule can only be determined by independent oversight bodies.
Such an ethics review should be mandatory for any analysis of hacked data and journals should require a disclosure statement about used data before publishing a study.

\section*{Evaluation of guidelines}

~\textcite{nature} propose a restrictive approach.
By default, hacked data should never be used in data science.
Only in rare cases, when several requirements are met, an exception can be made.
The first three requirements are prerequisites for using hacked data in data science, the following two about information obligations, followed by the requirement of independent oversight.

In my opinion, the first requirements, uniqueness, risk-benefit assessment, and consent, are very hard to fulfill.
It seems difficult to verifiably prove, that the hacked data could not be retrieved by any other way.
Additionally, it is questionable, whether it is ever possible to correctly assess the risk of the analysis of hacked data.
On top of that, the term ``minimal risk'' as well as ``benefits clearly outweighing harm'' leave much space for interpretation.
Furthermore, it seems most challenging to gather consent of all involved people or even debrief those about the usage of private hacked data.\\
I think the requirements regarding information obligations, traceability and privacy, are easier fulfill.
Regarding the illicit background of hacked data it is necessary and helpful to provide information on these topics when further processing the data.
As the required traceability and privacy information is mostly information on conducted procedures by a data scientist himself, it is the description of internal procedures without external dependencies on other parties.\\
Regarding the requirement of ethical oversight, the question arises whether ethical committees have the resources to externally assess those requirements for all publicly available hacked datasets and data scientists that want to use them.
I think it is very important that ethical committees assist data scientists in determining whether the conditions are fulfilled, but it cannot be their responsibility alone.

While I understand, that hacked data should only be used as last resort in an analysis, I think the proposed guidelines are only applicable to very rare cases of hacked data.
Additionally, despite the proposed, restrictive unethical-by-default rule and the guidelines, the usage of hacked data in data science itself cannot be stopped by them.
A data scientist is always able to process and analyze publicly available hacked data.
Only the publishing of his research in journals can be prevented in case a disclosure statement by an ethics committee is required.

Therefore, in my opinion, this whole discussion falls back on the establishment of a code of ethics for data science.
When data science perceives itself as profession and develops a code of ethics, guidelines regarding the use of hacked data can explicitly be included.
This would transfer proposed guidelines to fixed moral obligations every data scientist would follow and rise awareness for the topic.
In my opinion, this is the only way to sustain a lasting, cautious treatment of hacked data in data science.
